\Large\textbf{Advection versus diffusion in brain ventricular transport}

### 2025-04-11 paper/main.tex
conclusion and outlook

\subsection{Numerical considerations}
When simulating solute transport, we handled the advective transport
through the anterior/posterior boundaries
(where cardiac pressure pulsations were imposed)
via an approximate periodic boundary condition
that conserved the total solute concentration.
A possible improvement would be
to consider a 3D-0D coupling of the in- and outflow using
ordinary differential equations,
similar to applications in cardiac
modeling~\cite{Brown2024AMechanics, Augustin2021ACirculation}.
Alternatively, the beating cardiac motion could be modeled by ventricular wall deformation to avoid advective boundary fluxes\cite{Causemann2022HumanFramework,
Kurtcuoglu2005ComputationalSystem, Kurtcuoglu2007ComputationalSylvius, Linninger2005PulsatileBrain}.
%
We also remark that the Brezzi-Douglas-Marini and discontinuous Galerkin scheme
$\mathrm{BDM}_1$-$\mathrm{DG}_0$
used to discretize the Stokes equations
is computationally more expensive
than more commons discretizations such as 
Taylor-Hood ($P_2-P_{1}$ elements)~\cite{Stenberg1990ErrorProblem} or lower-order enriched
elements (e.g. MINI)~\cite{Brezzi2011MixedMethods}.
%Here, $P_k$ is the space of continuous Lagrange polynomials of degree less
%than or equal to the integer $k$.
However, we prefer the BDM-DG scheme because of its mass conservation properties: the numerical approximation of
the CSF velocity $\uu_h$ satisfies the continuity equation (\Cref{eq:stokes_eqs_cont}) point-wise, an important property for the stability of the advection-diffusion
equation, especially in advection-dominated cases.
\cite{cesmelioglu2022compatible, johnson2009numerical, gresho1998incompressible}
Additionally, the BDM-DG scheme 
lets us impose the impermeability condition $\uu\cdot\nn=0$
directly~\cite{Brezzi1985TwoProblems} (see \Cref{subsec:appendixA1} for more details).
%% For elements where this is not the case (such as Taylor-Hood elements),
%% the impermeability boundary condition must
%% be handled with for example a Nitsche method~\cite{Nitsche1971UberSind}
%% or with a Lagrange multiplier~\cite{Babuska1973TheMultipliers,
%% Bertoluzza2017BoundaryHemodynamics}, which affects
%% convergence properties \miro{$<<$what do you mean here?} and increases problem complexity
%% (see \Cref{subsec:appendixA1} for more details).

\subsection{Outlook}
The flow generated by ciliary motion
in zebrafish larval brain ventricles
is of the scale of the geometry~\cite{Olstad2019CiliaryDevelopment},
while in large brains such as the human brain, ciliary beating 
contributes to flow proximal to the ventricular walls
\cite{Siyahhan2014FlowVentricles}.
The cilia may play an important role in regulating the
amount of substances like extracellular vesicles close to the ventricular walls,
similar to coral reefs where cilia are primary regulators of
oxygen amounts at the coral surface~\cite{Pacherres2022CiliaryProduction}.
Cilia indeed regulate nodal flow at early
developmental stages of the brain~\cite{Hirokawa2006NodalAsymmetry},
and the discovery of a cilia-driven flow network
in rodent brains suggests such selective
brain ventricular transport~\cite{Faubel2016Cilia-basedVentricles}.
As the CSF flows through the central nervous system,
the system may be affected by the changes in CSF composition
due to reduced mixing following reduced
ciliary motion. We demonstrate that solute concentrations,
and in turn CSF composition, 
are affected by ventricular
geometry changes. A 10\% difference in volume
had a noticeable impact on solute transport,
showing that pathological ventricular volume changes,
such as the enlargement observed in hydrocephalus,
could affect CSF composition.
Because CSF plays an important role in the regulation of solutes
within the ventricular system~\cite{yamada2021cerebrospinal},
understanding more of the interplay between
transport dynamics and ventricular morphology,
as well as the impact of
reduced mixing resulting from cilia-deficiency,
are important considerations for future work.
%We hope our findings and the computational
%framework we have presented can help address
%such open questions in the future.

>>>>

\subsection{Numerical considerations}
When simulating solute transport, we handled the advective transport
through the anterior/posterior boundaries
(where cardiac pressure pulsations were imposed)
via an approximate periodic boundary condition
that conserved the total solute concentration.
A possible improvement would be
to consider a 3D-0D coupling of the in- and outflow using
ordinary differential equations,
similar to applications in cardiac
modeling~\cite{Brown2024AMechanics, Augustin2021ACirculation}.
Alternatively, the beating cardiac motion could be modeled by ventricular wall deformation to avoid advective boundary fluxes\cite{Causemann2022HumanFramework,
Kurtcuoglu2005ComputationalSystem, Kurtcuoglu2007ComputationalSylvius, Linninger2005PulsatileBrain}.
%
We also remark that the Brezzi-Douglas-Marini and discontinuous Galerkin scheme
$\mathrm{BDM}_1$-$\mathrm{DG}_0$
used to discretize the Stokes equations
is computationally more expensive
than more common discretizations such as 
Taylor-Hood ($P_2-P_{1}$ elements)~\cite{Stenberg1990ErrorProblem} or lower-order enriched
elements (e.g. MINI)~\cite{Brezzi2011MixedMethods}.
However, we prefer the BDM-DG scheme because of its mass conservation properties: the numerical approximation of
the CSF velocity $\uu_h$ satisfies the continuity equation (\Cref{eq:stokes_eqs_cont}) point-wise, an important property for the stability of the advection-diffusion
equation, especially in advection-dominated cases.
\cite{cesmelioglu2022compatible, johnson2009numerical, gresho1998incompressible}
Additionally, the BDM-DG scheme 
lets us impose the impermeability condition $\uu\cdot\nn=0$
directly~\cite{Brezzi1985TwoProblems} (see \Cref{subsec:appendixA1} for more details).

\subsection{Conclusion and outlook}
Altogether, we have shown that the flow
generated by ciliary motion in the zebrafish larval brain ventricles
is of the scale of the geometry~\cite{Olstad2019CiliaryDevelopment}.
In contrast, one can expect that in large brains such as the
human brain, ciliary beating contributes only to flow proximal
to the ventricular walls\cite{Siyahhan2014FlowVentricles}.
Nevertheless, the cilia may still play an important role in regulating the
amount of substances, especially extracellular vesicles,
close to the ventricular walls,
similar to coral reefs where cilia are primary regulators of
oxygen amounts at the coral surface~\cite{Pacherres2022CiliaryProduction}.
Notably, the discovery of a cilia-driven flow network
in rodent brains strongly supports such selective
brain ventricular transport~\cite{Faubel2016Cilia-basedVentricles}.
Our results also revealed that solute concentrations are affected
by ventricular geometry changes.
In fact, a 9\% difference in volume
had a noticeable impact on solute transport,
suggesting that pathological ventricular volume changes,
such as the enlargement observed in hydrocephalus,
could affect CSF distribution and composition.
Because CSF plays an important role in the regulation of solutes
within the ventricular system~\cite{yamada2021cerebrospinal},
understanding more of the interplay between
transport dynamics and ventricular morphology,
as well as the impact of
reduced mixing resulting from cilia-deficiency,
are important considerations for future work.
